On the number of tiles visited by a\\ line segment on a rectangular grid

### 2021-09-25 paper.tex
Confirmed new formula in Theorem 2 (\ref{eq: theo: funl, form; jsoli})
Proof of new formula in Theorem 2 (\ref{theo: funl, form})

\begin{figure}%
\centering%
\includegraphics[width=.7\textwidth]{mss}%
\caption{Obtaining $(\isolr, \jsolr)$ and $(i_\tiles, j_\tiles)$ in  Proposition~\ref{prop: min suff set, form}. Example with $a=1.35$, $b=1$, $\tiles=7$%
}%
\label{fig: min suff set, form}%
\end{figure}%

@@

\begin{theorem}
\label{theo: funl, form}
For $a, b \in \mathbb R^+$ and $\tiles \in \mathbb N$,
\begin{equation}
\label{eq: theo: funl, form; funl}
\funl(\tiles) = \sqrt{(\isoli-2)^2 a^2 + (\jsoli-2)^2 b^2}
\end{equation}
with
\begin{align}
\label{eq: theo: funl, form; isolr}
\isolr &= \max\left\{ \frac{(\tiles-3) b^2}{a^2+b^2}, 0 \right\} + 2, \\
\label{eq: theo: funl, form; jsolr}
\jsolr &= \max\left\{ \frac{(\tiles-3) a^2}{a^2+b^2}, 0 \right\} + 2, \\
\label{eq: theo: funl, form; isoli}
\isoli &= \left\lfloor \isolr + \frac 1 2 \right\rfloor, \\
\label{eq: theo: funl, form; jsoli}
\jsoli &= \left\lceil \jsolr - \frac 1 2 \right\rceil.
\end{align}
Equivalently, **I think this is valid for t geq 3 only, and needs a modification**
\begin{equation}
\label{eq: theo: funl, form; funl 2}
\funl(\tiles) = \sqrt{\frac{(t-3)^2 a^2 b^2}{a^2+b^2} + \roundterm^2(a^2+b^2)}
\end{equation}
with
\begin{equation}
\label{eq: roundterm}
\roundterm = |\isoli - \isolr| = |\jsoli - \jsolr|.
\end{equation}
This function is monotone increasing for $\tiles \geq 3$.
\end{theorem}

\begin{proof}
The $(i_\tiles,j_\tiles)$ pair in the minimal sufficient set $\mss$ corresponds to a maximum of $\tiles$ visited tiles. By construction of this set, any segment that visits $\tiles$ tiles must have length greater than $\sqrt{(i_\tiles-2)^2a^2 + (j_\tiles-2)^2b^2}$. For $\tiles \geq 3$ the expressions \eqref{eq: theo: funl, form; isoli} and \eqref{eq: theo: funl, form; jsoli} coincide with \eqref{eq: min suff set, form, i} and \eqref{eq: min suff set, form, j}. Therefore \eqref{eq: theo: funl, form; funl} gives the desired result. For $\tiles \in \{1, 2\}$ both \eqref{eq: theo: funl, form; isoli} and \eqref{eq: theo: funl, form; jsoli} equal $2$, and \eqref{eq: theo: funl, form; funl} gives $0$, which is the correct result.

***Proof of new formula. Perhaps it will reuse Figure~\ref{fig: min suff set, form}.

@@

***Comment on how the new formula has two terms with definite meanings: The first term is the distance-squared to the closest point on the diagonal, The second term is additional distance-squared incurred from rounding (i,j) to integers
***Figure for that? \ref{fig: min suff set, form} would do, but it's within a proof and uses different notation for the final point

>>>>

% ***This figure and Figure~\ref{fig: ijLS_1p35_detail} label the points as (i,j) instead of (ia,jb). Clarify in the text, or change the figures
\begin{figure}%
\centering%
\includegraphics[width=.7\textwidth]{mss}%
\caption{Obtaining $(\isolr, \jsolr)$ and $(i_\tiles, j_\tiles)$ in  Proposition~\ref{prop: min suff set, form}. Example with $a=1.35$, $b=1$, $\tiles=7$%
}%
\label{fig: min suff set, form}%
\end{figure}%

@@

\begin{theorem}
\label{theo: funl, form}
For $a, b \in \mathbb R^+$ and $\tiles \in \mathbb N$,
\begin{equation}
\label{eq: theo: funl, form; funl}
\funl(\tiles) = \sqrt{(\isoli-2)^2 a^2 + (\jsoli-2)^2 b^2}
\end{equation}
with
\begin{align}
\label{eq: theo: funl, form; isolr}
\isolr &= \max\left\{ \frac{(\tiles-3) b^2}{a^2+b^2}, 0 \right\} + 2, \\
\label{eq: theo: funl, form; jsolr}
\jsolr &= \max\left\{ \frac{(\tiles-3) a^2}{a^2+b^2}, 0 \right\} + 2, \\
\label{eq: theo: funl, form; isoli}
\isoli &= \left\lfloor \isolr + \frac 1 2 \right\rfloor, \\
\label{eq: theo: funl, form; jsoli}
\jsoli &= \left\lceil \jsolr - \frac 1 2 \right\rceil.
\end{align}
Equivalently, for $\tiles \geq 3$,
\begin{equation}
\label{eq: theo: funl, form; funl 2}
\funl(\tiles) = \sqrt{\frac{(t-3)^2 a^2 b^2}{a^2+b^2} + \roundterm^2(a^2+b^2)}
\end{equation}
with
\begin{equation}
\label{eq: roundterm}
\roundterm = |\isoli - \isolr| = |\jsoli - \jsolr|.
\end{equation}
This function is monotone increasing for $\tiles \geq 3$.
\end{theorem}

\begin{proof}
The $(i_\tiles,j_\tiles)$ pair in the minimal sufficient set $\mss$ corresponds to a maximum of $\tiles$ visited tiles. By construction of this set, any segment that visits $\tiles$ tiles must have length greater than $\sqrt{(i_\tiles-2)^2a^2 + (j_\tiles-2)^2b^2}$. For $\tiles \geq 3$ the variables $\isoli$, $\jsoli$ in \eqref{eq: theo: funl, form; isoli} and \eqref{eq: theo: funl, form; jsoli} respectively coincide with $i_\tiles$, $j_\tiles$ as given by \eqref{eq: min suff set, form, i} and \eqref{eq: min suff set, form, j}. Therefore \eqref{eq: theo: funl, form; funl} gives the correct result. For $\tiles \in \{1, 2\}$ both \eqref{eq: theo: funl, form; isoli} and \eqref{eq: theo: funl, form; jsoli} equal $2$, and \eqref{eq: theo: funl, form; funl} gives $0$, which is again the correct result.

For $\tiles \geq 3$, the term $\sqrt{(i_\tiles-2)^2a^2 + (j_\tiles-2)^2b^2}$ can be interpreted geometrically as the distance between $(i_\tiles, j_\tiles)$ and $(2a, 2b)$. As can be seen with the help of Figure~\ref{fig: min suff set, form}, this distance is the hypotenuse of a right triangle whose other two sides extend from $(2a, 2b)$ to $(\isolr a, \jsolr b)$ and from $(\isolr a, \jsolr b)$ to $(i_\tiles, j_\tiles)$ respectively. Therefore,
\begin{equation}
\label{eq: distance as hypotenuse}
\begin{split}
(i_\tiles-2)^2a^2 + (j_\tiles-2)^2b^2 &= (\isolr-2)^2a^2 + (\jsolr-2)^2b^2 \\
&\quad + (i_\tiles-\isolr)^2a^2 + (j_\tiles-\jsolr)^2b^2.
\end{split}
\end{equation}
For $\tiles \geq 3$ it stems from \eqref{eq: theo: funl, form; isolr}
and \eqref{eq: theo: funl, form; jsolr} that
\begin{equation}
\label{eq: distance as hypotenuse, first leg}
(\isolr-2)^2a^2 + (\jsolr-2)^2b^2 =
\frac{(\tiles-3)^2 (b^4a^2+a^2b^4)}{(a^2+b^2)^2} =
\frac{(\tiles-3)^2 a^2b^2}{a^2+b^2}.
\end{equation}
The fact that both $(\isolr, \jsolr)$ and $(i_\tiles, j_\tiles)$ are on the line $i+j-1=\tiles$ implies that $\isolr+\jsolr = i_\tiles+ j_\tiles$, and thus $|i_\tiles-\isolr| = |j_\tiles-\jsolr| = \roundterm$. Consequently,
\begin{equation}
\label{eq: distance as hypotenuse, second leg}
(i_\tiles-\isolr)^2a^2 + (j_\tiles-\jsolr)^2b^2 = \roundterm^2 (a^2+b^2).
\end{equation}
Substituting \eqref{eq: distance as hypotenuse, first leg} and \eqref{eq: distance as hypotenuse, second leg} into \eqref{eq: distance as hypotenuse} yields \eqref{eq: theo: funl, form; funl 2}.

@@

***Comment on how the new formula has two terms with definite meanings: The first term is the distance-squared to the closest point on the diagonal, The second term is additional distance-squared incurred from rounding (i,j) to integers
***I think we can do without a figure (Figure~\ref{fig: min suff set, form} is inside a proof and uses different notation for the final point).
